Title

### 2020-12-03 documents/DraftReport.Rmd
Added introduction

1. Briefly describe the issue, conclude with teaser sentence about what this project does

- IPMs as an increasingly popular tool that can provide tangible benefits  
-- Non-ID parameters, improved precision  
-- Expanded spatio-temporal inference  
-- These benefits have motivated interesting expansions and applications of this technique  
--- Examples   
- Brief history of IPM development  
-- First (?) joint analysis of mark-recovery and census data (Besbeas et al. 2002)  
-- Application to species (bats) w/ demographic data deficiencies (Schaub et al. 2007)  
-- Estimating hidden demographic processes (e.g., proportion of female lambs on the island and temporal variation) (Tavecchia et al. 2009)  
*“We experimentally evaluated the flexibility and limitations of the integrated approach by omitting portions of the data and comparing the results from these “incomplete” data sets with those obtained from the entire data set and from the analysis of the mark-resighting-recovery data only”*  
-- Inclusion of multivariate observations w/ estimation of transition probabilities within multi-state systems (McCrea et al. 2012)  
-- Inclusion of environmental stochasticity (Oppel et al. 2014)  
- However, this growth in usage has occurred with relatively poor understanding of how to assess goodness of fit, and therefore less is also known about the bounds/limits of the IPM framework to estimate demographic parameters in situations where available data might be sparse. 
- Study summary: In this study, we use simulated data to examine IPM performance across a range of scenarios where data informing key demographic parameters may be either sparse or missing.

2. With the expansion of the use of IPMs, few studies have robustly examined IPM performance through data simulation. 
- In those that have used data simulation, the focus has been on:
-- Proof of concept (improved precision and reduced bias): McCrea et al 2010, Chandler & Clark 2014, Schaub & Fletcher 2015, Bled et al. 2017, Ahrestani et al. 2017, Bowler et al. 2019  
- Model assumptions  
-- Independence assumption (Abadi et al. 2010, Schaub & Fletcher 2015, Weegman et al. 2020)  
-- MR assumptions (survival heterogeneity, marker-induced bias; Riecke et al. 2019)  
- None/few to date have looked at ability to return parameter values or power to detect trends when survival not informed by data or when a given data source is sparse, biased, or highly uncertain, but this is an interesting, complicated, and important topic  
- IPM performance is often context-specific: depends on life history, sample size, effect sizes, likelihood structures, study/inference goals, so what can we learn  
- Contributions of different datasets to the likelihood - helps us understand the underlying connections/mechanisms of the IPM, and to use this knowledge to think about IPM performance across life histories, survey designs, etc.

3. The “typical” IPM has gotten a lot of mileage, but can we push the limits?  
- The typical (survival and counts) has taught a lot, and been extended  
- Examples of a few cool extensions: spatially explicit IPMs, source-sink dynamics, environmental stochasticity
- But wouldn’t it also be cool and open doors if we could apply this approach in situations where this “typical” data wasn’t available or was sparse?  
-- Situations where only know fecundity and counts  
-- Situations where some of our data streams are biased and/or uncertain?  
-- Best to worst: unbiased/low variance, unbiased/high variance, biased/low variance, biased/high variance
--- Sun et al. 2019 (adding sites vs. changing detection)  
--- Simmonds et al. 2020; Farr et al. 2020 (similar to our goals but for SDMs)  
- We can’t confidently go in these directions without refining our understanding of the mechanisms underlying IPMs in these situations and where and when IPMs may not lead to improved precision or where they actually can’t estimate that unidentifiable parameter. This study fills that knowledge gap by X (restate purpose/approach) with the aim of improving our understanding of this important statistical approach, particularly for passerine bird life history.

4. Why this study is useful for practitioners    
- Study design: minimizing cost/effort and maximizing usefulness
- Surveys per week, hours per survey, number of sites, etc  
- Intentional survey design vs opportunistic/community science program data: Sanderlin et al. 2014
- Helping ecologists understand what is possible with available data, particularly given the tendency to maximize what you have 
- Importance of unbiased estimates for conservation and management

>>>>

Understanding the processes that control population dynamics in wild populations is a critical component of conservation but is frequently constrained by missing demographic information, uncertainty in data, and temporally or spatially misaligned datasets (citation). Integrated population models (IPMs) are an increasingly popular tool in ecology that offer release from these limitations by providing a framework to combine disparate datasets in a single, unified analysis (Schaub & Abadi 2011). This approach can provide several advantages like estimation of otherwise non-identifiable parameters, improved precision in parameter estimation, and [spatio-temporal inference] (citation). These advances have practical benefits for resource managers that must make decisions based on limited available information. As IPMs become more widely used, however, it is critical that they are robustly evaluated.

Since the introduction of IPMs to ecology in the early 2000s and following the work of Besbeas and colleagues with passerines (Besbeas et al. 2002, 2005), the common approach has been to estimate parameters by combining mark-recapture and abundance data into a joint likelihood. Researchers have has since expanded the use of IPMs to include a variety of taxa (e.g. in bats Rhinolophus ferrumequinum: Schaub et al. 2007, in Svalbard reindeer Rangifer tarandus platyrhynchus: Lee et al. 2015, Antarctic toothfish Dissostichus mawsoni: Mormede et al. 2019, wild boar Sus scrofa: Gamelon et al. 2019), [include multivariate observations with estimation of transition probabilities within multi-state and multi-site systems] (McCrea et al. 2012), evaluate management efforts (Cruz et al. 2018), and to include environmental processes and stochasticity (Oppel et al. 2014, Woodworth et al. 2017). This growth in usage has occurred despite a relatively poor understanding of how to assess goodness of fit however, and we lack knowledge of the limits of the IPM framework in situations where available data are sparse, of poor quality, or when certain data sources are absent. For example, many field studies may have survey and reproductive data while missing survival estimates from mark-recapture datasets.

Data simulation studies provide an avenue for robust evaluation of the IPM framework, but few such studies exist. Those that have used data simulation largely focus on proof of 1) concept (i.e. improved precision and reduced bias McCrea et al 2010, Chandler & Clark 2014, Schaub & Fletcher 2015, Bled et al. 2017, Ahrestani et al. 2017, Bowler et al. 2019), and 2) model assumptions such as independence of datasets (Abadi et al. 2010, Schaub & Fletcher 2015, Weegman et al. 2020). Riecke et al. (2019) found that parameter estimates from IPMs were highly sensitive to violations of mark-recapture model assumptions (i.e. marker-induced bias and survival heterogeneity). These findings further motivate studies to determine the reliability of IPMs in parameter estimation and in detecting trends when survival is not informed by data or when a data source is sparse, biased, or highly uncertain.

We aim to improve our understanding of this important statistical approach, particularly for passerine bird life history. In this study, we simulate data to examine IPM performance across a range of scenarios where data informing key demographic parameters are sparse or missing. We simulate three datasets for a for a theoretical passerine: 1) census counts, 2) reproductive output, and 3) mark-recapture (MR) of banded individuals. We then systematically vary the uncertainty in our abundance and MR data and run the integrated models under scenarios where demographic parameters like survival and fecundity are high or low. By identifying scenarios where IPMs do not lead to improved precision or cannot estimate unidentifiable parameters, we refine our understanding of the mechanisms underlying IPMs and can better identify when they may serve as a useful tool.
